---
title: "6373_Project_2"
output: html_document
editor_options: 
  chunk_output_type: inline
---



  The zip file includes two datasets: 
      1) bank-additional-full.csv with all examples, ordered by date (from May 2008 to November 2010).
      2) bank-additional.csv with 10% of the examples (4119), randomly selected from bank-additional-full.csv.
   The smallest dataset is provided to test more computationally demanding machine learning algorithms (e.g., SVM).

   The binary classification goal is to predict if the client will subscribe a bank term deposit (variable y).

5. Number of Instances: 41188 for bank-additional-full.csv

6. Number of Attributes: 20 + output attribute.

7. Attribute information:

   For more information, read [Moro et al., 2014].

   Input variables:
   # bank client data:
   1 - age (numeric)
   2 - job : type of job (categorical: "admin.","blue-collar","entrepreneur","housemaid","management","retired","self-employed","services","student","technician","unemployed","unknown")
   3 - marital : marital status (categorical: "divorced","married","single","unknown"; note: "divorced" means divorced or widowed)
   4 - education (categorical: "basic.4y","basic.6y","basic.9y","high.school","illiterate","professional.course","university.degree","unknown")
   5 - default: has credit in default? (categorical: "no","yes","unknown")
   6 - housing: has housing loan? (categorical: "no","yes","unknown")
   7 - loan: has personal loan? (categorical: "no","yes","unknown")
   # related with the last contact of the current campaign:
   8 - contact: contact communication type (categorical: "cellular","telephone") 
   9 - month: last contact month of year (categorical: "jan", "feb", "mar", ..., "nov", "dec")
  10 - day_of_week: last contact day of the week (categorical: "mon","tue","wed","thu","fri")
  11 - duration: last contact duration, in seconds (numeric). Important note:  this attribute highly affects the output target (e.g., if duration=0 then y="no"). Yet, the duration is not known before a call is performed. Also, after the end of the call y is obviously known. Thus, this input should only be included for benchmark purposes and should be discarded if the intention is to have a realistic predictive model.
   # other attributes:
  12 - campaign: number of contacts performed during this campaign and for this client (numeric, includes last contact)
  13 - pdays: number of days that passed by after the client was last contacted from a previous campaign (numeric; 999 means client was not previously contacted)
  14 - previous: number of contacts performed before this campaign and for this client (numeric)
  15 - poutcome: outcome of the previous marketing campaign (categorical: "failure","nonexistent","success")
   # social and economic context attributes
  16 - emp.var.rate: employment variation rate - quarterly indicator (numeric)
  17 - cons.price.idx: consumer price index - monthly indicator (numeric)     
  18 - cons.conf.idx: consumer confidence index - monthly indicator (numeric)     
  19 - euribor3m: euribor 3 month rate - daily indicator (numeric)
  20 - nr.employed: number of employees - quarterly indicator (numeric)

  Output variable (desired target):
  21 - y - has the client subscribed a term deposit? (binary: "yes","no")

8. Missing Attribute Values: There are several missing values in some categorical attributes, all coded with the "unknown" label. These missing values can be treated as a possible class label or using deletion or imputation techniques. 

```{r Import Bank Additional Full data set}
library(tidyverse)
library(ggplot2)
library(GGally)
library(MASS)
library(mvtnorm)
library(ROSE)

bankdata <- read.csv("bank-additional-full.csv", header=TRUE, sep = ';')
summary(bankdata)


#Train/Test set split
bank_balance <- ovun.sample(y~., data=bankdata, method="under",seed= 1234)$data
table(bank_balance$y)

set.seed(1234)
splitPerc = .85
trainIndices = sample(1:dim(bank_balance)[1], round(splitPerc * dim(bank_balance)[1]))
train = bank_balance[trainIndices,]
test = bank_balance[-trainIndices,]

```

```{r GGpairs Variables 1-7}
ggpairs(bankdata,columns = 1:7,aes(color=y),progress = FALSE)

```

```{r GGpairs Variables 8-11}
ggpairs(bankdata,columns = 8:11,aes(color=y),progress=FALSE)
```

```{r GGpairs Variables 12-20}
ggpairs(bankdata,columns = 12:20,aes(color=y), progress=FALSE)
```

```{r Conduct PCA in prep of LDA}
bankcont <- bankdata[c(1,11:14,16:20)]








summary(bankcont_train)
bankcont_test <- test[c(1,11:14,16:20)]
pc.test <-prcomp(bankcont_test,scale.=TRUE)
pc.test <-data.frame(pc.test$x)
pc.test$y <-test$y

sum(diag(cov(bankcont_train)))

pc.result<-prcomp(bankcont_train,scale.=TRUE)
pc.scores<-pc.result$x
pc.result$rotation

par(mfrow=c(1,2))
eigenvals<-(pc.result$sdev)^2
plot(1:10,eigenvals/sum(eigenvals),type="l",main="Scree Plot",ylab="Prop. Var. Explained")
cumulative.prop<-cumsum(eigenvals/sum(eigenvals))
plot(1:10,cumulative.prop,type="l",main="Cumulative proportion",ylim=c(0,1))
par(mfrow=c(1,1))

pc.scores<-data.frame(pc.scores)
pc.scores$y <-train$y

ggplot(data=pc.scores, aes(x=PC1,y=PC2)) +geom_point(aes(col=y), size=1)+ggtitle("PCA of Opening account")
ggplot(data=pc.scores, aes(x=PC1,y=PC3)) +geom_point(aes(col=y), size=1)+ggtitle("PCA of Opening account")
ggplot(data=pc.scores, aes(x=PC1,y=PC4)) +geom_point(aes(col=y), size=1)+ggtitle("PCA of Opening account")
ggplot(data=pc.scores, aes(x=PC2,y=PC3)) +geom_point(aes(col=y), size=1)+ggtitle("PCA of Opening account")
ggplot(data=pc.scores, aes(x=PC2,y=PC4)) +geom_point(aes(col=y), size=1)+ggtitle("PCA of Opening account")
ggplot(data=pc.scores, aes(x=PC3,y=PC4)) +geom_point(aes(col=y), size=1)+ggtitle("PCA of Opening account")
```

```{r Perform LDA}
mylda <- lda(y ~ PC1+PC2+PC3+PC4, data = pc.scores)
pred<- predict(mylda,newdata=pc.test)$class

Truth <-pc.test$y
x<-table(pred,Truth)
x

```
