---
title: "Project_2_EDA"
author: "Satvik Ajmera"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

The data is related with direct marketing campaigns (phone calls) of a Portuguese banking institution. The classification goal is to predict if the client will subscribe a term deposit (variable y).

  The zip file includes two datasets: 
      1) bank-additional-full.csv with all examples, ordered by date (from May 2008 to November 2010).
      2) bank-additional.csv with 10% of the examples (4119), randomly selected from bank-additional-full.csv.
   The smallest dataset is provided to test more computationally demanding machine learning algorithms (e.g., SVM).

   The binary classification goal is to predict if the client will subscribe a bank term deposit (variable y).

5. Number of Instances: 41188 for bank-additional-full.csv

6. Number of Attributes: 20 + output attribute.

7. Attribute information:

   For more information, read [Moro et al., 2014].

   Input variables:
   # bank client data:
   1 - age (numeric)
   2 - job : type of job (categorical: "admin.","blue-collar","entrepreneur","housemaid","management","retired","self-employed","services","student","technician","unemployed","unknown")
   3 - marital : marital status (categorical: "divorced","married","single","unknown"; note: "divorced" means divorced or widowed)
   4 - education (categorical: "basic.4y","basic.6y","basic.9y","high.school","illiterate","professional.course","university.degree","unknown")
   5 - default: has credit in default? (categorical: "no","yes","unknown")
   6 - housing: has housing loan? (categorical: "no","yes","unknown")
   7 - loan: has personal loan? (categorical: "no","yes","unknown")
   # related with the last contact of the current campaign:
   8 - contact: contact communication type (categorical: "cellular","telephone") 
   9 - month: last contact month of year (categorical: "jan", "feb", "mar", ..., "nov", "dec")
  10 - day_of_week: last contact day of the week (categorical: "mon","tue","wed","thu","fri")
  11 - duration: last contact duration, in seconds (numeric). Important note:  this attribute highly affects the output target (e.g., if duration=0 then y="no"). Yet, the duration is not known before a call is performed. Also, after the end of the call y is obviously known. Thus, this input should only be included for benchmark purposes and should be discarded if the intention is to have a realistic predictive model.
   # other attributes:
  12 - campaign: number of contacts performed during this campaign and for this client (numeric, includes last contact)
  13 - pdays: number of days that passed by after the client was last contacted from a previous campaign (numeric; 999 means client was not previously contacted)
  14 - previous: number of contacts performed before this campaign and for this client (numeric)
  15 - poutcome: outcome of the previous marketing campaign (categorical: "failure","nonexistent","success")
   # social and economic context attributes
  16 - emp.var.rate: employment variation rate - quarterly indicator (numeric)
  17 - cons.price.idx: consumer price index - monthly indicator (numeric)     
  18 - cons.conf.idx: consumer confidence index - monthly indicator (numeric)     
  19 - euribor3m: euribor 3 month rate - daily indicator (numeric)
  20 - nr.employed: number of employees - quarterly indicator (numeric)

  Output variable (desired target):
  21 - y - has the client subscribed a term deposit? (binary: "yes","no")

8. Missing Attribute Values: There are several missing values in some categorical attributes, all coded with the "unknown" label. These missing values can be treated as a possible class label or using deletion or imputation techniques. 


```{r}
library(tidyverse)
library(ggplot2)
library(GGally)
#bank_full.path <- file.choose()
#hello = read.csv(bank_full.path, header = TRUE, sep = ";")
bank_full <- read.csv("bank-additional-full.csv", header=TRUE, sep = ';')
summary(bank_full)
sapply(bank_full, class)


#Convert Columns to Numeric
#Move previous into other list 
num.cols <- c("age","duration","campaign","pdays","previous","emp.var.rate","cons.price.idx","cons.conf.idx","euribor3m","nr.employed")
bank_full[num.cols] <- lapply(bank_full[num.cols], as.numeric)
sapply(bank_full, class)

#Convert Columns to Factor
#Note: Converted previous to factor
fac.col <- c("job","marital","education","default","housing","loan","contact","month","day_of_week", "poutcome","y","previous")
bank_full[fac.col] <- lapply(bank_full[fac.col], as.factor)
sapply(bank_full, class)
conv_bank = read.csv("bank-additional-full.csv", header=TRUE, sep = ';',stringsAsFactors = F)
#Convert Columns to Factor
check = names(conv_bank)
conv_bank[check] <- lapply(conv_bank[check], factor)
sapply(conv_bank, class)
num.cols <- c("age","duration","campaign","pdays","previous","emp.var.rate","cons.price.idx","cons.conf.idx","euribor3m","nr.employed")
conv_bank[num.cols] <- lapply(conv_bank[num.cols], as.numeric)
sapply(conv_bank, class)


head(bank_full)
colnames(bank_full)
bank_full
summary(bank_full)
```
response (y)
has the client subscribed a term deposit?
no: 36548
yes: 4640

```{r}
attach(bank_full)

t(aggregate(y~duration,data=bank_full,summary))
plot(y~duration,col=c("red","blue"))
```

```{r}
#Proportion Tables
prop.table(table(y,job),2)
prop.table(table(y,age),2)
prop.table(table(y,marital),2)
prop.table(table(y,education),2)
prop.table(table(y,duration),2)
prop.table(table(y,campaign),2)
prop.table(table(y,housing),2)
prop.table(table(y,contact),2)
prop.table(table(y,day),2)
prop.table(table(y,month),2)
prop.table(table(y,poutcome),2)
prop.table(table(y,pdays),2)
prop.table(table(y,previous),2)
prop.table(table(y,loan),2)

#Proportion plots
plot(y~job,col=c("red","blue"))
plot(y~age,col=c("red","blue"))
plot(y~marital,col=c("red","blue"))
plot(y~education,col=c("red","blue"))
plot(y~duration,col=c("red","blue"))
plot(y~marital,col=c("red","blue"))
plot(y~campaign,col=c("red","blue"))
plot(y~housing,col=c("red","blue"))
plot(y~contact,col=c("red","blue"))
plot(y~day,col=c("red","blue"))
plot(y~month,col=c("red","blue"))
plot(y~poutcome,col=c("red","blue"))
plot(y~pdays,col=c("red","blue"))
plot(y~previous,col=c("red","blue"))
plot(y~loan,col=c("red","blue"))
```

```{r}
#Histograms by Response Variable Y
ggplot(data = bank_full, aes(x = age,fill = y)) + geom_histogram()+facet_wrap(~y,
scales = "free_y")
ggplot(data = bank_full, aes(x = age,fill = y)) + geom_histogram()+facet_wrap(~y,
scales = "free_y")
ggplot(data = bank_full, aes(x = duration,fill = y)) + geom_histogram()+facet_wrap(~y,
scales = "free_y")
ggplot(data = bank_full, aes(x = duration,fill = y)) + geom_histogram()+facet_wrap(~y,
scales = "free_y")
ggplot(data = bank_full, aes(x = campaign,fill = y)) + geom_histogram()+facet_wrap(~y,
scales = "free_y")
ggplot(data = bank_full, aes(x = emp.var.rate,fill = y)) + geom_histogram()+facet_wrap(~y,
scales = "free_y")
ggplot(data = bank_full, aes(x = emp.var.rate)) + geom_histogram()
t(aggregate(y~age,data=bank_full,summary))
```

```{r}
#More Proportion Plots
ggplot(data = bank_full, aes(x = job,fill=job)) + geom_bar()+ggtitle("Job Type Count")
#Job Proportion Plot
ggplot(bank_full,aes(x = job,fill = y)) + geom_bar(position = "fill") + ggtitle("Proportion of Response for Job Type")
#Education
ggplot(data = bank_full, aes(x = education,fill=education)) + geom_bar()+ggtitle("Education Type Count")
#Education Proportion Plot
ggplot(bank_full,aes(x = education,fill = y)) + geom_bar(position = "fill") + 
ggtitle("Proportion of Response for Education")
ggplot(data = bank_full, aes(x = duration, fill = age)) + geom_histogram(fill="blue",bins = 20)
```


```{r}
cat.newBankAdd <- bank_full[-c(2:10,15)]
colnames(newBankAdd)
cols <- c("age","duration","campaign","pdays","previous","emp.var.rate","cons.price.idx","cons.conf.idx","euribor3m","nr.employed")
cat.newBankAdd[cols] <- lapply(cat.newBankAdd[cols], factor)
sapply(cat.newBankAdd, class)
cat.newBankAdd
plot(cat.newBankAdd$y~cat.newBankAdd$previous,col=c("red","blue"))

#Education Proportion Plot
ggplot(cat.newBankAdd,aes(x = previous,fill = y)) + geom_bar(position = "fill")
#All the social economic variables are very related. So I either drop them entirely or keep one?
ggplot(cat.newBankAdd,aes(x = emp.var.rate,fill = y)) + geom_bar(position = "fill")
ggplot(cat.newBankAdd,aes(x = cons.price.idx,fill = y)) + geom_bar(position = "fill")
ggplot(cat.newBankAdd,aes(x = euribor3m,fill = y)) + geom_bar(position = "fill")
#Age is definitely correlated with Yes, significantly lower aged and higher aged people said yes.
ggplot(cat.newBankAdd,aes(x = age,fill = y)) + geom_bar(position = "fill")
ggplot(cat.newBankAdd,aes(x = duration,fill = y)) + geom_bar(position = "fill")
ggplot(bank_full,aes(x = month,fill = y)) + geom_bar(position = "fill") + ggtitle("Proportion for Month")
ggplot(bank_full,aes(x = day_of_week,fill = y)) + geom_bar(position = "fill")

prop.table(table(cat.newBankAdd$y,cat.newBankAdd$duration),2)
#distribution of continuous variables
#Previous is problematic because it should be a categorized as a categorical variable. it has 11 levels.

summary(cat.newBankAdd)

t(aggregate(euribor3m~y,data=bank_full,summary))
t(aggregate(emp.var.rate~y,data=bank_full,summary))
t(aggregate(previous~y,data=cat.newBankAdd,summary))
t(aggregate(duration~y,data=bank_full,summary))
# prop.table(table(,))
ftable(addmargins(table(bank_full$duration,bank_full$y)))
prop.table(bank_full$y,bank_full$duration)
```


```{r}
library(ggplot2)
library(gplots)
#removed categorical variables for PCA and checking multicolinearity
newBankAdd <- bank_full[-c(2:10,15)]

View(newBankAdd)
summary(newBankAdd)
colnames(newBankAdd)

#Correlations between continuous variables
my.cor<-cor(newBankAdd[,1:10])
my.cor
#Correlation Heatmap
heatmap.2(my.cor,col=redgreen(75), 
          density.info="none", trace="none", dendrogram=c("row"), 
          symm=F,symkey=T,symbreaks=T, scale="none")

mr.cor2 = cor(pc.result)

```


```{r}
#Removed Previous because it is a categorical variable and the seperation looks significantly better.
#Now we only need 4 PCs to explain 80% of the variance.
newBankAdd <- bank_full[-c(2:10,15)]
newBankAdd <- newBankAdd[-c(5)]
ggpairs(newBankAdd, aes(colour = y, alpha = 0.2), lower=list(combo=wrap("facethist",  
binwidth=0.5)))
head(newBankAdd)
newBankAdd.test3 <- newBankAdd[-c(3,6)]
pc.result<-prcomp(newBankAdd[,1:9],scale.=TRUE)
pc.scores<-pc.result$x
pc.scores<-data.frame(pc.scores)
pc.scores$y<-newBankAdd$y
# test=cor(pc.result$x)
# cor(pc.result$rotation)
# table = pc.result$rotation
#Use ggplot2 to plot the first few pc's
ggplot(data = pc.scores, aes(x = PC1, y = PC2)) +
  geom_point(aes(col=y), size=.5)+
  ggtitle("PCA of Bank Full Additional")

ggplot(data = pc.scores, aes(x = PC2, y = PC3)) +
  geom_point(aes(col=y), size=.5)+
  ggtitle("PCA of Bank Full Additional")

ggplot(data = pc.scores, aes(x = PC3, y = PC4)) +
  geom_point(aes(col=y), size=.5)+
  ggtitle("PCA of Bank Full Additional")

ggplot(data = pc.scores, aes(x = PC4, y = PC5)) +
  geom_point(aes(col=y), size=.5)+
  ggtitle("PCA of Bank Full Additional")

ggplot(data = pc.scores, aes(x = PC1, y = PC9)) +
  geom_point(aes(col=y), size=.5)+
  ggtitle("PCA of Bank Full Additional")
#Scree Plot
eigenvals<-(pc.result$sdev)^2
plot(1:9,eigenvals/sum(eigenvals),type="l",main="Scree Plot PC's",ylab="Prop. Var. Explained",ylim=c(0,1))
cumulative.prop<-cumsum(eigenvals/sum(eigenvals))
lines(1:9,cumulative.prop,lty=2)
plot(pc.scores[,3], pc.scores[,4])
plot(pc.scores)
library(GGally)
ggpairs(pc.result$x[,2:5])

```


```{r}
library(ResourceSelection)
library(car)
bank_full_omit <-na.omit(bank_full)
o.newBankAdd <-na.omit(newBankAdd)
#Only continuous Predictors fit
con.bankmodel<-glm(y ~ ., data=o.newBankAdd,family = binomial)
alias(con.bankmodel)
(vif(con.bankmodel))


#Full model fit
full.bankmodel<-glm(y ~ ., data=bank_full_omit,family = binomial)
alias(full.bankmodel)
summary(bankmodel)

alias(bankmodel)
#Using this tool, GVIF is the same as VIF for continuous predictors only
#For categorical predictors, the value GVIG^(1/(2*df)) should be squared and interpreted
#as a usuaul vif type metric.The following code can be used to interpret VIFs like we 
#discussed in class.
(vif(con.bankmodel))
(vif(full.bankmodel))
vif(bankmodel)[,]
par(mfrow = c(2, 2))
plot(model.main)
```

```{r}
t(aggregate(month~y,data=bank_full,summary))
t(aggregate(marital~y,data=bank_full,summary))
prop.table(table(y,marital),2)

#looks like there is a higher proportion of no's than yes to some of the types of jobs.
t(aggregate(job~y,data=bank_full,summary))
prop.table(table(y,job),2)

t(aggregate(education~y,data=bank_full,summary))
prop.table(table(y,education),2)

#contact by cellular yield a higher percent of yes compared to telephone
t(aggregate(contact~y,data=bank_full,summary))
prop.table(table(y,contact),2)

```



